---
title: "df"
author: "B198096"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
## Set up the environment 
## And specify all of the parameters 

```{r, initializing}
library(DoubletFinder)
#https://github.com/chris-mcginnis-ucsf/DoubletFinder
library(Seurat)
library(tidyverse)
library(ggplot2)

sce_list <- list()
files <- list()

range <- 1:9
name <- "ramachandran_fibrosis_cell_"
label <- "rf"
dir <- "~/Desktop/Bioinformatics/Dissertation/ramachandran_seurat/"
files <- paste0(dir,name,range,"_QC.rds")
unfiltered_counts <- readRDS(file = "~/Desktop/Bioinformatics/Dissertation/ramachandran_seurat/rf_unfiltered_dim")


for (i in range){
  sce_list[i] <- readRDS(file = files[[i]])
} 
```

## Calculate the expected doublets based on the QC_unfiltered counts (returned cells) 
## This is based on 10x genomics's reference table for expected doublets based on returned cells 

```{r, expected doublet calc}
exp_d <- list()
for (i in range){
  #Total cells recovered 
  cells <- unfiltered_counts[[i]]
  #10x calcualtes expected doublets as a function of number of recovered cells 
  #Their table is for brackets of 1000 cells, so I am setting the boundries at 500 cells 
  #So that the bracket used is whatever 1000s is closest 
  bracket <- as.integer(cells/1000)
  bracket_up <- cells%%1000
  #Then this is a conditional to assign expected doublet rate   
  #Based on the bracket information 
  if (cells < 500) {
    exp_d[[i]] <- 0.004
    } else {
      exp_d[[i]] <- 0.008 * bracket
      if (bracket_up > 500) {
        exp_d[[i]] <- 0.008 * (bracket + 1)
      }
    }
  }
```

## Then run DoubletFinder
## The loop runs through and simulates pK values and then saves the optimal value
## Then run DoubletFinder using this pK value and the expected doublets from above 

```{r}
pK <- list()
nExp_poi <- list()
nExp_poi.adj <- list()
df_col <- list()
for (i in range){
  #This simulates various pK values for this data to identify the optimal value 
  #I am specifying that 30 PCs were used in the original PCA (I could change 
  #This for the DF analysis, but choosing to use all PCs)
  #And specifying that I used SCTransform 
  sweep.res.list_tf1 <- paramSweep_v3(sce_list[[i]], PCs = 1:30, sct = TRUE)
  #Generates a summary table of the above simultions. 
  #No multiplexed data so GT is false 
  sweep.stats_tf1 <- summarizeSweep(sweep.res.list_tf1, GT = FALSE)
  #Then this calculates BC metrics as a score of success for each pK value 
  bcmvn_tf1 <- find.pK(sweep.stats_tf1)
  #This goes through and pulls the optimal pK value after the simulations 
  pK_temp <- bcmvn_tf1 %>%
    filter(BCmetric == max(BCmetric)) %>%
    select(pK)
  pK[[i]] <- as.numeric(as.character(pK_temp[[1]]))
  #Adding the annotations of cluster results from the PCA 
  annotations <- sce_list[[i]]@meta.data$ClusteringResults
  #Modelling homotypic doublets based on clustering patterns 
  homotypic.prop <- modelHomotypic(annotations)
  #Predicting the number of expected doublets based on exp_d from above 
  nExp_poi[[i]] <- round(exp_d[[i]]*nrow(sce_list[[i]]@meta.data))  
  #Adjusting the total expected doublet count
  nExp_poi.adj[[i]] <- round(nExp_poi[[i]]*(1-homotypic.prop))
  #Running DoubletFinder 
  #I am setting PCs to 30 again 
  #The developers found that results are not significantly impacted by pN 
  #And so I followed their default value of 0.25 
  #pK and nExp are calculated above 
  #reuse.pANN is not being used, and SCTransform was used
  sce_list[[i]] <- doubletFinder_v3(sce_list[[i]], PCs = 1:30, pN = 0.25, pK = pK[[i]], nExp = nExp_poi[[i]], reuse.pANN = FALSE, sct = TRUE)
  #sce_list[[i]] <- doubletFinder_v3(sce_list[[i]], PCs = 1:30, pN = 0.25, pK = pK[[i]], nExp = nExp_poi.adj[[i]], reuse.pANN = paste0("pANN_0.25_",pK[[i]],"_",nExp_poi[[i]]), sct = FALSE)
  df_col[[i]] <- paste0("DF.classifications_0.25_",pK[[i]],"_",nExp_poi[[i]])
  DimPlot(sce_list[[i]], reduction = 'umap', group.by = df_col[[i]])
}

for (i in range){
  saveRDS(sce_list[[i]], file = paste0(dir,name,i,"_DF.rds"))
  }
```
## Including Plots
## Save all of the labeled doublet plots 
```{r}
for (i in range){
  plt <- DimPlot(sce_list[[i]], reduction = 'umap', group.by = df_col[[i]])
  ggsave(filename = paste0(dir,label,i,"_df_plot.pdf"), plot = plt)
  plt
}
```

## Save the pK, nExp_poi, and df_col values in case we need them later or for the report 
```{r pressure, echo=FALSE}
saveRDS(pK, file = paste0(dir,label,"_pKs.rds"))
saveRDS(nExp_poi, file = paste0(dir,label,"_nExp_poi.rds"))
saveRDS(df_col, file = paste0(dir,label,"_df_col.rds"))
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
